---
title: "Strings and Factors"
output: github_document
---


```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```


## Strings and manipulations
detecting patterns in string vectors
string matches remember are case sensitive
use str_detect to see if something exists; often useful with filter function
str_replace
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")

str_detect(string_vec, "Jeff")

str_replace(string_vec, "jeff", "Jeff")

```

looking for expressions that start with I think; ^ indicates beginning of the line; $ end of a line
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")

str_detect(string_vec, "i think$")

```

look for capital or lower case pumpkin; specify two diff cases for a match
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")
```


starts with a number then has any lower case letter that follows that 

have to be explicit about allowing spaces; spaces count as a character
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9]")

str_detect(string_vec, "^[0-9]")

str_detect(string_vec, "^[0-9][a-z]")

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

. matches any character at all
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")

```

find any instance where the square bracket exists; square bracket is a special character so you designate wtih two backslashes
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")

str_detect(string_vec, "\\[[0-9]")

```

try not to use special characters in your data bc it's very annoying!! stick with letters, numbers, and some underscores

## Factors

create a string variable
```{r}
vec_sex = c("male", "male", "female", "female")
vec_sex
```

convert to factor; print out to see the factor levels and their order; as numeric to see undelrying numbers
```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)

vec_sex = fct_relevel(vec_sex, "male")
vec_sex

as.numeric(vec_sex)
```

if you have categorical variables don't label as 1-5; put actual informative labels for the categories 
working with factor variables be sure you know what is going on behind the scenes

## NSDUH
```{r}

```






